Estimating real-time highstreet footfall from Wi-Fi probe requests

### 2018-06-24 paper/paper.tex
wed end of the day
writing from bus
journal template and more writing

\documentclass[11t, a4paper, twocolumn]{article} 
\input{structure.tex}
\addbibresource{references.bib}
\title{Estimating real-time highstreet footfall from Wi-Fi probe requests}
\author{
	\authorstyle{
		Balamurugan Soundararaj\textsuperscript{1}, 
		James Cheshire\textsuperscript{1} and 
		Paul Longley\textsuperscript{1}}
	\newline\newline
	\textsuperscript{1} \institution{
		Department of Geography, 
		University College London, 
		United Kingdom } }
\date{\today}
\begin{document}
	\maketitle
	\thispagestyle{firstpage}
	\textbf{Abstract -}
		\input{sections/abstract}
	\section{Introduction}\label{introduction}
		\input{sections/introduction}
	\section{Previous Work}\label{previous_work}
		\input{sections/previous_work}
	\section{Methodology}\label{methodology}
		\input{sections/methodology}
	\section{Pilot Study}\label{pilot_study}
		\input{sections/pilot_study}
	\section{Main Study}\label{main_study}
		\input{sections/main_study}
	\section{Discussion}\label{discussion}
		\input{sections/discussion}
	\section{Conclusion}\label{conclusion}
		\input{sections/discussion}
	\printbibliography[title={References}]
\end{document}

>>>>

\documentclass[]{interact}
\usepackage{epstopdf}
\usepackage{subfigure}
\usepackage{natbib}
\bibpunct[, ]{(}{)}{,}{a}{}{,}
\renewcommand\bibfont{\fontsize{10}{12}\selectfont}
\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}
\theoremstyle{remark}
\newtheorem{remark}{Remark}
\newtheorem{notation}{Notation}

\begin{document}

	\title{Estimating real-time highstreet footfall from Wi-Fi probe requests}
	\author{
		\name{ Balamurugan Soundararaj\textsuperscript{a}\thanks{CONTACT Balamurugan Soundararaj. Email: s.bala@ucl.ac.uk}, James Cheshire\textsuperscript{a} and Paul Longley\textsuperscript{a} }
		\affil{\textsuperscript{a}Department of Geography, University College London, United Kingdom.}
	}
	\maketitle

	\begin{abstract}
		\input{sections/abstract} \end{abstract}
	\begin{keywords}
		Highstreet footfall; Wi-Fi Probe requests; Sensors; MAC Randomisation \end{keywords}
	\section{Introduction}\label{introduction}
		\input{sections/introduction}
	\section{Previous Work}\label{previous_work}
		\input{sections/previous_work}
	\section{Methodology}\label{methodology}
		\input{sections/methodology}
	\section{Pilot Study}\label{pilot_study}
		\input{sections/pilot_study}
	\section{Main Study}\label{main_study}
		\input{sections/main_study}
	\section{Conclusion}\label{conclusion}
		\input{sections/conclusion}
	\section*{Acknowledgement}
		\input{sections/acknowledgement}
	\bibliographystyle{tfv} \bibliography{references}

\end{document}

### 2018-06-24 paper/sections/introduction.tex
journal template and more writing

Probe requests are low level signals standardised by IEEE 802.1b/g specification as the first step in establishing a Wi-Fi based connection between two devices and is implemented in any Wi-Fi capable device irrespective of the manufacturer or the model. This ubiquity and standardisation make them an excellent source of open, passive, continuous, and wireless data generated by Wi-Fi capable devices present at any given time and location. Considering the unprecendented levels of mobile device ownership in recent years, we can in turn use this data to understand the population distribution in highly dynamic urban environments with high spatial and temporal granularity \citep{freud2015,konto2017}.

>>>>

Probe requests are low level signals standardised by IEEE 802.1b/g specification \citep{ieee2013} as the first step in establishing a Wi-Fi based connection between two devices and is implemented in any Wi-Fi capable device irrespective of the manufacturer or the model. This ubiquity and standardisation make them an excellent source of open, passive, continuous, and wireless data generated by Wi-Fi capable devices present at any given time and location. Considering the unprecendented levels of mobile device ownership in recent years, we can in turn use this data to understand the population distribution in highly dynamic urban environments with high spatial and temporal granularity \citep{freud2015,konto2017}.

### 2018-06-24 paper/sections/main_study.tex
main analysis
journal template and more writing

The aims of the main study are,
test the validity of the signal strength algorithm in different 
micro site conditions.
Test that the sequence number algorithm works in real world
for different locations and different times.
Test if the calibration works over intervals
Finally conclude if we can estimate footfall confidently
with just probe requests.

\subsection{locations}
five locations were selected across central london which
had different types of configurations and specific problems
configuratons are shown in fig. map is shown in figure.
\begin{enumerate}
	\item Phone Shop Camden - has phones and bus stops.
	\item Restaurant TCR - has seating area on either side.
	\item Holborn Information Kiosk - High volumne station entry
	\item Restaurant Russell Square - seating on one side and side walk on other
	\item Shop Charring Cross - sidewalk on one side and phone shop next door
\end{enumerate}
installations were carried out over the time period from xxxx to xxxx.
the data collection happened from xxxx to xxxx. Manual counting was carried out
with high precision on dates xxxx and aggregated five minutes on xxxx.
The difference in methods could lead to some inaccuracies in data.
The overall statistics of data collected. 
The overally schedule is shown in half page graphic.

\subsection{Signal strength filtering}

\subsection{Device Fingerprinting}

\subsection{Manual Calibration}

\subsection{Discussion}

>>>>

The aims of the main study are,
test the validity of the signal strength algorithm in different 
micro site conditions.
Test that the sequence number algorithm works in real world
for different locations and different times.
check if the thresholds are consistent.
Test if the calibration works over intervals
Finally conclude if we can estimate footfall confidently
with just probe requests.

five locations were selected across central london which
had different types of configurations and specific problems
configuratons are shown in fig. map is shown in figure.
\begin{enumerate}
	\item Phone Shop Camden - has phones and bus stops.
	\item Restaurant TCR - has seating area on either side.
	\item Holborn Information Kiosk - High volumne station entry
	\item Restaurant Russell Square - seating on one side and side walk on other
	\item Shop Charring Cross - sidewalk on one side and phone shop next door
\end{enumerate}
Installations were carried out over the time period from xxxx to xxxx.
the data collection happened from xxxx to xxxx. Manual counting was carried out
with high precision on dates xxxx and aggregated five minutes on xxxx.
The difference in methods could lead to some inaccuracies in data.
The overall schedule is shown in half page graphic.

The overall statistics of data collected. How many probe requests.
Comparision of location in terms of volume, patterns in daily footfall etc.
The comparision between global and local.
comparision between different types of vendors.
Specifics on top 5 manufacturers.

We do a daily analysis of distribution of signal strengths.
The thresholds are shown in the table.
The average is xxxx and standard deviation is xxx.
we notice that the variation is lot. There is a definite
change with the micro site locations.

We see how the signal strength filtering affects the counts.
compared to manual counts we look at the average mean errors (daily)
per 5 minutes. The counts go as follows.
Shown as a red line in the Figure.

We can conclude that even though it has variations, this is
a good method to reduce the overall level of error.

This is also done for different locations hourly for all the data we had.
we compare it to the manual counts and see that the average mean error
has been reduced/increased. The finger print works well for all the locations.
It also works over a period of time and gives us a comparable and close 
footfall count to the manual count. The thresholds found in the pilot study
works as well.

Finally we normalise the sensor counts to match the manual count using
a fraction/ adjustment factor calculated from the know manual counts.
we have three sets of counts. We check if the adjustment factor holds the
same in all three counts across locations. It does with a variations from
xxx to xxx. The results are shown in the table.

we see that the signal strength filtering works and reduces error by
xxxxx. there is variation by locations.
we see that sequence number algorithm works as well. The threshold stays
constant well and works well across locations. The calibration also works
and ajustment factor stays consistent short term. This needs more work 
long term.

### 2018-06-24 paper/sections/methodology.tex
journal template and more writing

The overall methodology is to use a WiFi antenna to collect probe requests,
assign a unique identification based on the device that created the request,
aggregate them based on the unique identification for a specific time interval
to create an estimate of number of people at the location.
In this section we look at the characteristics of probe requests in detail,
outline the methodology used to collect these probe requests, 
look at the uncertainities and biases in the process and 
device methods to overcome these issues.

\subsection{Probe Requests}
Probe request is a low level packet sent by a mobile device as a means of
"scanning" for various access points available at a specific location.
This packet consists information about the mobile device
including but not limited to,
Media Access Control address - A two part 12 bit identifier where the first part
identifies the manufacturer of the device and the second part identifies the
device itself. This MAC address is of two types, global - the real identifier
of the device which doesnt change and local - virtual, random addresses used
for temporary situations.
Sequence number of the packet to keep track of the replies.
The access points for which the packet is being sent to.
Capabilities of the device.
We can also infer other things about the packet such as
time at which the packet has been received,
total length of the packet
time it took to transmit the packet
Signal strength of the packet.
All of this can help us identifying the packets and label them when they are transmitted by the same device.

@@

The sensor consists of wireless transponder in a linux/ MacOSX based machine
which is put into a monitor mode and
wireshark software (tshark) is used to sniff, filter and
save the relevant packets.
The filters applied are management frame > probe requests > broadcast. 
This is then parsed, personal information is hashed, 
compressed and sent to a server via web socket protocol.
The server receives, logs and stores the data in a postgresql database,
for further analysis.
The manual count was undertaken using an Android application
on a mobile phone where timestamp of every individual pedestrian footfall
was recorded by touch screen. A schematic is shown in ref figure.

>>>>

The primary aim of this research is to enable us to collect a series of probe requests and process them into an usable pedestrian footfall count. We do this by using a Wi-Fi receiver to collect probe requests broadcasted by mobile devices, filtering out the background noise and aggregating them based on the device that generated them. We begin by looking at the characteristics of probe requests in detail, device a methodology to collect these probe requests in public areas, examine the systemic biases and uncertainties in the data collection method and device data processing methods to overcome these challanges. Finally we compare the processed footfall counts to the ground truth recorded by primary surveys.

\subsection{Probe Requests}
Probe requests are a special type of management packets broadcast by Wi-Fi enabled devices as part of the various functions such as scanning for available access points (AP), quick geo-location by triangulation based known APs, etc.
These are broadcast by all Wi-Fi enabled devices regardless of the manufacturer, type or model of the devices though there is some variation on the frequency and the information transmitted through them.
In some cases, such as Android devices, these are broadcast even when the Wi-Fi functionality has been turned off by the user.
Thus these signals can be used to reliably identify the presence of Wi-Fi enabled mobile devices.

Being a first step of connection initiated by the mobile device, these packets have information regarding the characteristics of the mobile device itself. Some of the key information we can infer from these requests are,
\begin{enumerate}
	\item Media Access Control (MAC) address which is an unique identifier for the wireless hardware of the mobile device,
	\item Sequence number of the request for the mobile device to keep track of the responses,
	\item Timestamp at which the request was received by the AP,
	\item Total length of the request in number of bits, and 
	\item The strength of the signal which transmitted the request.
\end{enumerate}
The MAC address is the primary unique identifier for the mobile device.
It has two parts, first part gives information about the manufacturer of the device and the second is uniue to the device. The MAC address can be randomised (hence non unique) and is marked as such. Though sequence number and length of the packet are not strictly unique, we hypothesize that we can use them to estimate unique devices.

@@

The sensor consists of wireless transponder in a linux/ MacOSX based machine
which is put into a monitor mode and
wireshark software (tshark) is used to sniff, filter and
save the relevant packets.
The filters applied are management frame to probe requests to broadcast. 
This is then parsed, personal information is hashed, 
compressed and sent to a server via web socket protocol.
The server receives, logs and stores the data in a postgresql database,
for further analysis.
The manual count was undertaken using an Android application
on a mobile phone where timestamp of every individual pedestrian footfall
was recorded by touch screen. A schematic is shown in ref figure.
